---
title: "HW2"
author: "Lin Wang"
date: "9/19/2020"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(stats)
```
## Data Wrangling

1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r read and merge data}
dat1 <- read.csv("C:/Users/Hope/Documents/PM566HW/chs_individual.txt")
dim(dat1)

dat2 <- read.csv("C:/Users/Hope/Documents/PM566HW/chs_regional.txt")
dim(dat2)

dat <- merge(
  x = dat1, y = dat2, by.x = "townname", by.y = "townname", all.x = TRUE, all.y = FALSE
) 

# take a look at data after merging
head(dat)
summary(dat)
# dealing with NAs
dat <- as.data.table(dat)

dat[, bmi := fcoalesce(bmi, mean(bmi, na.rm = TRUE)),
    by = .(male, hispanic)]

dat[, fev := fcoalesce(fev, mean(fev, na.rm = TRUE)),
    by = .(male, hispanic)]
```

- There are 1200 rows and 49 columns in dat with some missing values.

2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r create "obesity_level"}
# create variable
dat[bmi<14, obesity_level := "underweight"]
dat[bmi>=14 & bmi<22, obesity_level := "normal"]
dat[bmi>=22 & bmi<24, obesity_level := "overweight"]
dat[bmi>24, obesity_level := "obese"]

# summary table of obesity level
n <- count(dat, obesity_level)

merge(n, dat[, .(
  bmi_min = min(bmi, na.rm = TRUE),
  bmi_max = max(bmi, na.rm = TRUE)
),
by = obesity_level])
```

- The obesity_level is coded successfully. 

3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r create "smoke_gas_exposure"}
# create variable
dat[smoke == 0 & gasstove == 0, smoke_gas_exposure := "none_exposed"]
dat[smoke == 0 & gasstove == 1, smoke_gas_exposure := "gas_exposed"]
dat[smoke == 1 & gasstove == 0, smoke_gas_exposure := "smoke_exposed"]
dat[smoke == 1 & gasstove == 1, smoke_gas_exposure := "all_exposed"]

# dropping NAs
# dat <- dat[!is.na(smoke_gas_exposure)]

# summary table of smoke_gas_exposure
count(dat, smoke_gas_exposure)
```

- The smoke_gas_exposure variable has 4 categories which reflects the exposure degree to smoke and gas.

4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r create summary tables}
# FEV & asthma by town
unique(dat[, .(
  fev_avg = mean(fev, na.rm = TRUE),
  fev_sd = sd(fev, na.rm = TRUE)
  ),
  by = townname][order(townname)])

round(proportions(table(dat$townname, dat$asthma)), digits = 3)

# FEV & asthma by sex
unique(dat[, .(
  fev_avg = mean(fev, na.rm = TRUE),
  fev_sd = sd(fev, na.rm = TRUE)
  ),
  by = male][order(male)])

round(proportions(table(dat$male, dat$asthma)), digits = 3)

# FEV & asthma by obesity level
unique(dat[, .(
  fev_avg = mean(fev, na.rm = TRUE),
  fev_sd = sd(fev, na.rm = TRUE)
  ),
  by = obesity_level][order(obesity_level)])

round(proportions(table(dat$obesity_level, dat$asthma)), digits = 3)

# FEV & asthma by smoke_gas_exposure
unique(dat[, .(
  fev_avg = mean(fev, na.rm = TRUE),
  fev_sd = sd(fev, na.rm = TRUE)
  ),
  by = smoke_gas_exposure][order(smoke_gas_exposure)])

round(proportions(table(dat$smoke_gas_exposure, dat$asthma)), digits = 3)
```

- Here the average and sd of fev are successfully summarized, however, I'm not sure how to calculate the proportion of asthma(which is a binary variable) using data.table, so I simply use table() to generate the proportions.

## EDA
The primary questions:
- What is the association between BMI and FEV?
- What is the association between smoke and gas exposure and FEV?
- What is the association between PM2.5 exposure and FEV?

```{r EDA}
library(ggplot2)
```

1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r scatterplot}
ggplot(data = dat, 
       mapping = aes(x = bmi, y = fev, color = townname))+
  geom_point()+
  stat_smooth(method = lm)+
  facet_wrap(~townname, nrow = 4)
```

- FEV is positively associated with BMI, it becomes higher when BMI is higher.

2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r histogram}
ggplot(dat, mapping = aes(x = fev, color = obesity_level, fill = obesity_level))+
  geom_histogram(binwidth = 50, alpha = 0.7,position="identity")+
  scale_fill_brewer(palette = "PuOr")+
  scale_color_brewer(palette = "Purples")

dat %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(mapping = aes(x = fev, color = smoke_gas_exposure, fill = smoke_gas_exposure))+
  geom_histogram(binwidth = 50, position="identity")+
  scale_fill_brewer(palette = "PuOr")+
  scale_color_brewer(palette = "Purples")
```

- FEV is normally distributed under normal category, but is skewed under other categories, which shows FEV distribution is affected by obesity level.
- FEV is normally distributed by smoke_gas_exposure.

3. Barchart of BMI by smoke/gas exposure.

```{r}
dat %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
  ggplot()+
  geom_bar(mapping = aes(x = obesity_level, color = smoke_gas_exposure, fill = smoke_gas_exposure))+
  scale_fill_brewer(palette = "Purples")+
  scale_color_brewer(palette = "Purples")
```

- BMI is not associated with smoke or gas exposure.

4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
dat %>%
  filter(!(obesity_level %in% NA)) %>%
ggplot(mapping = aes(x = obesity_level, y = fev))+
  stat_summary(fun.data = "mean_sdl")

dat %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(mapping = aes(x = smoke_gas_exposure, y = fev))+
  stat_summary(fun.data = "mean_sdl")
```

- Mean FEV varies among obesity level categories, FEV is positively associated with BMI.
- Mean FEV is similar among smoke_gas_exposure categories, FEV is not associated with smoke or gas exposure.

5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
library(leaflet)

pm25.pal = colorNumeric(c("blue", "purple", "red"), domain = dat$pm25_mass)

leaflet(dat) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircles(lat = ~lat, lng = ~lon, opacity = 1, fillOpacity = 1, radius = 500, color = ~pm25.pal(pm25_mass)) %>%
  addLegend(position = "bottomleft", pal = pm25.pal, values = dat$pm25_mass , title = "PM2.5 mass concentrations", opacity = 1)
```

- We could see PM2.5 has higher mass concentrations around Los Angeles.

6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r PM2.5 ~ FEV}
dat %>%
  filter(!(pm25_mass %in% NA)) %>%
ggplot(mapping = aes(x = pm25_mass, y = fev))+
  geom_point()+
  stat_smooth(method = lm)+
  theme_classic()
```

- PM2.5 is not associated with FEV.

