---
title: "HW3"
author: "Lin Wang"
date: "10/10/2020"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(include  = TRUE)
library(httr)
library(xml2)
library(stringr)
```

## APIs

1. Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r retrieve paper, eval=TRUE, cache=TRUE}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")
counts
# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

- There are 560 papers related the term "sars-cov-2 trial vaccine".

2. Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r download paper details, eval=TRUE, }
# Find paper ids under the term
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db="pubmed",
    term="sars-cov-2 trial vaccine",
    retmax=1000)
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>") # "</?Id>"

# Now we have 225 ids
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db="pubmed",
    id=paste(ids, collapse=","),
    retmax=1000,
    rettype="abstract"
    )
)
# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


3. As we did in lab 7. Create a dataset containing the following: Pubmed ID number, Title of the paper, Name of the journal where it was published, Publication date, and Abstract of the paper (if any).

The `xml2::xml_children()` function to keep one element per id. This way, if a paper is missing the abstract, or something else, we will be able to properly match PUBMED IDS with their corresponding records.

```{r one-string-per-response, eval = TRUE}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

Now, extract the title, journal name, publication date and abstract

```{r extract details, eval=TRUE}
# get titles
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")

# get journal names
journals <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journals <- str_remove_all(journals, "</?[[:alnum:]]+>")
journals <- str_replace_all(journals, "\\s+", " ")

# get publication date
dates <- str_extract_all(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]]+>")
dates <- str_replace_all(dates, "\\s+", " ")

# get abstracts, check if any NA's
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+", " ")
table(is.na(abstracts))
```

- There are 61 papers missing abstracts.

```{r create a database, eval=TRUE}
database <- data.frame(
  PubMedID = ids,
  Title = titles,
  Journal = journals,
  Date = dates,
  Abstract = abstracts
)
knitr::kable(database)
```

- The dataset is successfully created.

## Text Mining

```{r}

```


