---
title: "HW3"
author: "Lin Wang"
date: "10/10/2020"
output: github_document
---

You can view the report [here](https://ghcdn.rawgit.org/HopeW233/Pm566HW/master/HW3/README.html)

```{r setup}
knitr::opts_chunk$set(include  = TRUE)
library(httr)
library(xml2)
library(stringr)
library(readr)
library(tidytext)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## APIs

1. Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r retrieve paper, eval=TRUE, cache=TRUE}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")
counts
# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

- There are 560 papers related the term "sars-cov-2 trial vaccine"(10/13/2020).

2. Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r download paper details, eval=TRUE}
# Find paper ids under the term
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db="pubmed",
    term="sars-cov-2 trial vaccine",
    retmax=1000)
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>") # "</?Id>"

# Now we have 225 ids
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db="pubmed",
    id=paste(ids, collapse=","),
    retmax=1000,
    rettype="abstract"
    )
)
# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


3. As we did in lab 7. Create a dataset containing the following: Pubmed ID number, Title of the paper, Name of the journal where it was published, Publication date, and Abstract of the paper (if any).

The `xml2::xml_children()` function to keep one element per id. This way, if a paper is missing the abstract, or something else, we will be able to properly match PUBMED IDS with their corresponding records.

```{r one-string-per-response, eval = TRUE}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

Now, extract the title, journal name, publication date and abstract

```{r extract details, eval=TRUE}
# get titles
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")

# get journal names
journals <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journals <- str_remove_all(journals, "</?[[:alnum:]]+>")
journals <- str_replace_all(journals, "\\s+", " ")

# get publication date
dates <- str_extract_all(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]]+>")
dates <- str_replace_all(dates, "\\s+", " ")

# get abstracts, check if any NA's
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+", " ")
table(is.na(abstracts))
```

- There are 61 papers missing abstracts.

```{r create a database, eval=TRUE}
database <- data.frame(
  PubMedID = ids,
  Title = titles,
  Journal = journals,
  Date = dates,
  Abstract = abstracts
)
knitr::kable(database[1:5, ])
```

- Check the first 5 papers and the dataset is successfully created.

## Text Mining

```{r get data, warning=FALSE}
dat <- read_csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")
```

1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stop words?

```{r count tokens}
# counting the number of each token
dat %>%
  unnest_tokens(token, abstract) %>%
  count(token, sort = TRUE) %>%
  top_n(10, n)

# removing stop words
dat %>%
  unnest_tokens(token, abstract) %>%
  filter(!(token %in% stop_words$word)) %>%
  count(token, sort = TRUE) %>%
  top_n(10, n)

# find 5 most common tokens
tokens <- dat %>%
  unnest_tokens(token, abstract) %>%
  filter(!(token %in% stop_words$word)) %>%
  group_by(term) %>%
  count(token, sort = TRUE) %>%
  top_n(5, n) %>%
  arrange(term, desc(n))

knitr::kable(tokens, caption = "Table 1: The 5 most common tokens for each search term")
```

- The top 10 tokens mostly are stop words, covid, 19 is the most frequent words.
- After removing stop words, covid, 19 is still the most frequent words, but some terms like cancer, prostate, eclampsia come up.
- The most 5 common tokens for each term are as table shows.

2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r bigrams}
dat %>%
  unnest_ngrams(token, abstract, n=2) %>%
  count(token, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(token, n)))+
  geom_col()+
  ggtitle("The top 10 common bigrams")

# removing stop words
dat_unitted <- dat %>%
  unnest_ngrams(token, abstract, n = 2) %>%
  separate(token, into = c("word1", "word2"), sep = " ") %>%
  anti_join(tidytext::stop_words, by = c("word1" = "word")) %>%
  anti_join(tidytext::stop_words, by = c("word2" = "word")) %>%
  unite(bigram, word1, word2, sep = " ")

dat_unitted %>%
  count(bigram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(x = n, y = fct_reorder(bigram, n)))+
  geom_col()+
  ggtitle("The top 10 common bigrams after removing stop words")
```

- After removing stop words, we could see top 10 common bigrams contain more about search terms such as covid 19, prostate cancer, pre eclampsia, systic fibrosis.

3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r TF-IDF}
tf_idf <- dat %>%
  unnest_tokens(token, abstract) %>%
  filter(!(token %in% stop_words$word)) %>%
  group_by(term) %>%
  count(token, sort = TRUE) %>%
  bind_tf_idf(token, term, n) %>%
  top_n(5, n) %>%
  arrange(term, desc(tf_idf)) 

knitr::kable(list(tf_idf[, c(1:3,6)], tokens), 
            caption = "Table 2: Top 5 tokens with the highest TF-IDF from each term VS Table 1: The 5 most common tokens for each search term")
```

- Under the term covid: covid and pandemic has the highest TF-IDF value, which excludes 19, patients and disease(TF-IDF=0).
- Under the term cystic fibrosis: cf becomes the first term instead of fibrosis.
- Under the term meningitis: patients has a TF-IDF value at 0.
- Under the term preeclampsia: tokens like pre and woman are excluded(TF-IDF=0).
- Under the term prostate cancer: only prostate has TF-IDF value.
